Contextual recruitment of selective attention can be updated via changes in task-relevance

### 2019-10-17 manuscript.tex
draft complete

\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}
\usepackage{setspace}\doublespacing
\newenvironment{zeroindent} {\par\setlength{\parindent}{0pt}} {\par}
\setlength{\parskip}{0pt}
\raggedbottom

@@

\abstract{
Evidence across a wide variety of attention paradigms shows that environmental cues can trigger adjustments to ongoing priorities for attending to relevant and irrelevant information. This context-specific control over attention suggests that cognitive control can be both automatic and flexible. For instance, in selective attention tasks, congruency effects are larger for items that appear in a context associated with infrequent conflict than in a context associated with frequent conflict. Since the to-be-presented context cannot be predicted or prepared for in advance, attention is assumed to be rapidly updated on-the-fly, triggered by the currently presented context. Context-specific control exemplifies how learning and memory processes can influence attention to enable cognitive flexibility. However, what determines the use of previously learned associations still remains unclear. In the current study, we examined whether task-relevance would influence the learning and use of context cues in a flanker task. Using a secondary counting task, context dimensions associated with differing levels of conflict were made task-relevant or -irrelevant across the experiment. In short, we found that making new contextual information task-relevant caused participants to ignore a previously learned context-attention association and adopt a new context-specific control strategy; all without changing the experimental stimuli. This result suggests that task-relevance is a key determinant of context-specific control.

@@

Selective attention is commonly investigated using interference paradigms such as the Stroop (1935) and flanker (Eriksen \& Eriksen, 1974) tasks, where participants identify a target while ignoring a response-congruent or -incongruent distractor. Performance is typically better on congruent versus incongruent trials and the difference---the congruency effect---taken as an index of attentional priorities. Large congruency effects are thought to reflect ineffective filtering of the distracting stimuli whereas small congruency effects are thought to reflect effective filtering. By probing factors that systematically alter congruency effects, we can then make inferences about processes that control attentional filtering. For example, manipulating the frequency of conflict via the proportion of congruent versus incongruent trials has shown to influence the size of the congruency effect. Typically, a high proportion congruent experiment produces large congruency effects, whereas a low proportion congruent experiment produces small congruency effects (Logan \& Zbrodoff, 1979; Lowe \& Mitterer, 1982; West \& Baylis, 1998). This result is usually explained as strategic control, where participants increase attentional control under high-conflict demands and relax attentional control under low-conflict demands (Logan, 1980; Logan \& Zbrodoff, 1979; Logan, Zbrodoff, \& Williamson, 1984; Lowe \& Mitterer, 1982). Recent work however, has demonstrated that attentional control is not only adjusted by top-down regulation, but can also be triggered automatically by environmental cues (Brosowsky \& Crump, 2018; Bugg \& Crump, 2012; Egner, 2014; Fischer \& Dreisbach, 2015; King, Korb, \& Egner, 2012; Mayr \& Bryck, 2007).

For example, Crump, Gong, and Milliken (Crump, Gong, \& Milliken, 2006; see also, Corballis \& Gratton, 2003) presented Stroop stimuli in one of two randomly chosen locations and manipulated the frequency of conflict associated with each location. One location was associated with a high frequency of conflict (25\% congruent trials) and the other with a low frequency of conflict (75\% congruent trials). Overall, the proportion of congruent trials was 50\% and randomized such that the upcoming location could not be predicted. Even so, congruency effects were shown to be smaller for trials where the stimulus appeared in the high conflict location as compared to the low conflict location. This effect, known now as the context-specific proportion congruent effect (CSPC), has now been replicated in a number of different selective attention paradigms (e.g., Alards-Tomalin, Brosowsky, \& Mondor, 2017; Blais, Harris, Sinanian, \& Bunge, 2015; Bugg, 2014; Crump, 2016; Crump, Milliken, Leboe-McGowan, Leboe-McGowan, \& Gao, 2018; Fischer, Gottschalk, \& Dreisbach, 2014; Hübner \& Mishra, 2016).

@@

One explanation for CSPC effects is that the repeated application of attentional priorities in a particular context creates an associative link in episodic memory between the attentional control procedures and the contextual information (Abrahamse, Braem, Notebaert, \& Verguts, 2016; Brosowsky \& Crump, 2018; Crump, 2016; Egner, 2014). Once the associative link is established, processing the context is assumed to trigger the retrieval of previous experiences, automatically reinstating associated attentional priorities. As a result, automatic memory-based retrieval allows attentional priorities to be rapidly adjusted in a context-appropriate fashion. However, the obligatory nature of memory-based attentional control remains unclear, in particular in situations when multiple contextual features could be used as cues for retrieval.

For example, real-world environments offer a plethora of context cues. Consider writing an article in a coffee shop. The coffee shop is a complex environment with numerous context cues (e.g., pictures on the wall, chairs, coffee, laptop), and associated attentional priorities. Although a laptop might cue attentional processing helpful for writing, other features of the coffee shop might cue attentional processing related to other experiences like socializing and people-watching. Given that all of these cues are available to set attentional priorities, what determines which associative relationships will be used to guide attention? On the one hand, context-based retrieval of attentional priorities might be completely obligatory and a writer's attention would be at the mercy of the environment. On the other hand, there is some evidence that contextual cuing of attention is not entirely obligatory and can be constrained by task goals, allowing the laptop to take priority as a contextual cue over other contextual cues that are irrelevant to the task at hand.

Prior work shows that task-relevance is necessary for establishing and using context-attention associations. First, the most commonly used context cue is location (Brosowsky \& Crump, 2016; Corballis \& Gratton, 2003; Crump, 2016; Crump et al., 2006; Crump, Vaquero, \& Milliken, 2008; Weidler \& Bugg, 2016; Weidler et al., 2018). Although in some sense, the location is irrelevant for identifying the color of a word, it seems disingenuous to consider it completely task-irrelevant. In order to identify the target, the participant must first identify and orient to its location on the screen. In this is sense, it is relevant to their ability to complete the task and furthermore, location information may receive priority during encoding (Logan, 1998; Mayr, 1996). Similar arguments could be made about other context features that are inherent to processing the target stimulus, like font-type (Bugg, Jacoby, \& Toth, 2008) and color (Vietze \& Wendt, 2009).

Second, we are aware of only one study that has looked at task-relevance within the CSPC literature. Crump, Vaquero, \& Milliken (2008) used a Stroop prime-probe task where the distractor word is presented first followed by a color patch probe. Critically, they used the shape of the color patch probe, either a square or a circle, as the contextual dimension; one associated with a high frequency of conflict, the other with a low frequency. In one experiment participants were made aware of the shape-proportion contingencies but given no additional instructions. In the second, participants were given a secondary task to count the number of trials that contained a square. The first experiment failed to find evidence for context-specificity, whereas the second did. Although not testing task-relevance directly, Cañadas, Rodríguez-Bailón, Milliken, and Lupiáñez (2013){]} found similar effects. In this study they used images of male and female faces as context cues in a flanker task. When given instructions to think of the faces as members of the gender categories they found CSPC effects, but when given instructions to think of the faces as individuals they did not. These results suggest that making a context dimension relevant to the on-going task is important for producing CSPC effects.

\hypertarget{the-current-study}{%
\section{The Current Study}\label{the-current-study}}

Prior work suggests that the task-relevance of contextual information is important for establishing new associations between context cues and attentional priorities. Simply put, CSPC effects emerge when context cues are somehow made relevant to the task, but do not when made irrelevant (Cañadas et al., 2013; Crump et al., 2006, 2008). However, it still remains unclear how task-relevance may or may not inform the use of previously learned associations. In the current study, we examined whether manipulating task-relevance would cause participants to ignore a previously learned association and adopt a new context-specific control strategy.

@@

The critical measure of interest is the congruency effect produced in the frequency unbiased context. If, on the one hand, task-relevance has no impact on the use of context-attention associations we would expect no differences between the congruency effects when the high-conflict context dimension is made task-relevant or the low-conflict context dimension is made task-relevant. This could occur because all three contexts are treated as individual contexts throughout the whole experiment (e.g., Cañadas et al., 2013). Or it could occur because associations formed in the first block interfere with learning new associations in the second (e.g., Brosowsky \& Crump, 2016). On other hand, task-relevance might not only dictate how associations are formed, but also whether learned associations are used. In this case, we would expect smaller congruency effects when the high-conflict context dimension is made task-relevant as compared to when the low-conflict context dimension is made task-relevant. This would demonstrate that participants were able to ignore a previously learned association and adopt a new context-specific control strategy.

@@

Experiment 1 used a 2x2x3 mixed design with task-relevant context (0\% and 100\%) and unbiased-item congruency (congruent and incongruent) as within-subject factors, and context-type (object, social, and social/non-repeating) as the between-subjects factor.
All three conditions were constructed using the same general method. The experiment was divided into two phases. Each phase consisted of 144 flanker trials (48 trials per context), and 13 count response trials for a total of 314 trials. On the count response trials, participants indicated how many trials they had counted until that point. The count response trials occurred once for every 12 flanker trials and was randomly inserted between trial 6 and 12 of each 12-trial block. Each phase ended with one additional count response trial.

@@

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/figure2-1.pdf}
\caption{\label{fig:figure2}Results from experiment 1 showing congruency effects (incongruent - congruent) for frequency unbiased items in reaction times (left) and error rates (right) as a function of the task-relevant context (100\% proportion congruent versus 0\% proportion congruent).}
\end{figure}

@@

The primary question of interest was whether the task-relevance of context features associated with different levels of conflict would influence the size of the congruency effect for frequency unbiased items (see Table 1). To that end, mean correct RTs and mean error rates from frequency unbiased trials were submitted to a mixed analysis of variance (ANOVA) with task-relevant context (0\% and 100\% PC) and unbiased-item congruency (congruent and incongruent) as within-subject factors, and context-type (object, social, and social/non-repeating) as the between-subjects factor.

The results of the RT analysis revealed the critical, two-way interaction between task-relevant context and unbiased-item congruency to be significant, \(F(1, 128) = 8.30\), \(\mathit{MSE} = 1,608.54\), \(p = .005\), \(\hat{\eta}^2_p = .061\), demonstrating smaller congruency effects when the context dimension associated with high conflict was made task-relevant. Additionally, the three-way interaction between the task-relevant context, unbiased-item congruency, and context-type was non-significant, \(F(2, 128) = 0.00\), \(\mathit{MSE} = 1,608.54\), \(p = .996\), \(\hat{\eta}^2_p = .000\), showing no evidence for differences between the CSPC effects across context-type manipulations.

Similarly, the corresponding error analysis also resulted in a significant two-way interaction between the task-relevant context and unbiased-item congruency, \(F(1, 128) = 6.72\), \(\mathit{MSE} = 13.93\), \(p = .011\), \(\hat{\eta}^2_p = .050\), and a non-significant three-way interaction between task-relevant context, unbiased-item congruency, and context-type, \(F(2, 128) = 0.92\), \(\mathit{MSE} = 13.93\), \(p = .400\), \(\hat{\eta}^2_p = .014\). Therefore, the congruency effects in error rates were also smaller when the context dimension associated with high conflict was made task-relevant, corroborating the results of the reaction time analysis.

>>>>

\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}
\usepackage{threeparttable}
\usepackage{setspace}

\singlespacing
\setlength{\parskip}{0pt}
\raggedbottom

@@

\abstract{
Evidence across a wide variety of attention paradigms shows that environmental cues can trigger adjustments to ongoing priorities for attending to relevant and irrelevant information. This context-specific control over attention suggests that cognitive control can be both automatic and flexible. For instance, in selective attention tasks, congruency effects are larger for items that appear in a context associated with infrequent conflict than in a context associated with frequent conflict. Since the to-be-presented context cannot be predicted or prepared for in advance, attention is assumed to be rapidly updated on-the-fly, triggered by the currently presented context. Context-specific control exemplifies how learning and memory processes can influence attention to enable cognitive flexibility. However, what determines the use of previously learned associations still remains unclear. In the current study, we examined whether task-relevance would influence the learning and use of context cues in a flanker task. Using a secondary counting task, context dimensions associated with differing levels of conflict were made task-relevant or -irrelevant across the experiment. In short, we found that making new contextual information task-relevant caused participants to suppress a previously learned context-attention association and adopt a new context-specific control strategy; all without changing the experimental stimuli. Furthermore, we found participants did not spontaneously learn about context-specific proportion manipulations and explicit instructions were insufficient for producing context-specific effects. These results suggests that task-relevance is a key determinant of context-specific control.

@@

Selective attention is commonly investigated using interference paradigms like the Stroop (1935) and flanker (Eriksen \& Eriksen, 1974) tasks, where participants identify a target while ignoring a response-congruent or -incongruent distractor. Performance is typically better on congruent versus incongruent trials and the difference---the congruency effect---taken as an index of attentional priorities. Large congruency effects are thought to reflect ineffective filtering of the distracting stimuli whereas small congruency effects are thought to reflect effective filtering. By probing factors that systematically alter congruency effects, we can then make inferences about processes that control attentional filtering. For example, manipulating the frequency of conflict via the proportion of congruent versus incongruent trials has shown to influence the size of the congruency effect. Typically, a high proportion congruent experiment produces large congruency effects, whereas a low proportion congruent experiment produces small congruency effects (Logan \& Zbrodoff, 1979; Lowe \& Mitterer, 1982; West \& Baylis, 1998). This result is usually explained as strategic control, where participants increase attentional control under high conflict demands and relax attentional control under low conflict demands (Logan, 1980; Logan \& Zbrodoff, 1979; Logan, Zbrodoff, \& Williamson, 1984; Lowe \& Mitterer, 1982). Recent work however, has demonstrated that attentional control is not only adjusted by top-down regulation, but can also be triggered automatically by environmental cues (Brosowsky \& Crump, 2018; Bugg \& Crump, 2012; Egner, 2014; Fischer \& Dreisbach, 2015; King, Korb, \& Egner, 2012; Mayr \& Bryck, 2007).

For example, Crump, Gong, and Milliken (Crump, Gong, \& Milliken, 2006; see also, Corballis \& Gratton, 2003) presented Stroop stimuli in one of two randomly chosen locations and manipulated the frequency of conflict associated with each location. One location was associated with a high frequency of conflict (25\% congruent trials) and the other with a low frequency of conflict (75\% congruent trials). Overall, the proportion of congruent trials was 50\% and trials were randomized such that the upcoming location could not be predicted. Even so, congruency effects were shown to be smaller for trials where the stimulus appeared in the high conflict location as compared to the low conflict location. This effect, known now as the context-specific proportion congruent effect (CSPC), has now been replicated in a number of different selective attention paradigms (e.g., Alards-Tomalin, Brosowsky, \& Mondor, 2017; Blais, Harris, Sinanian, \& Bunge, 2015; Bugg, 2014; Crump, 2016; Crump, Milliken, Leboe-McGowan, Leboe-McGowan, \& Gao, 2018; Fischer, Gottschalk, \& Dreisbach, 2014; Hübner \& Mishra, 2016).

@@

One explanation for CSPC effects is that the repeated application of attentional priorities in a particular context creates an associative link in episodic memory between the attentional control procedures and the contextual information (Abrahamse, Braem, Notebaert, \& Verguts, 2016; Brosowsky \& Crump, 2018; Crump, 2016; Egner, 2014). Once the associative link is established, processing the context is assumed to trigger the retrieval of previous experiences, automatically reinstating associated attentional priorities. As a result, automatic memory-based retrieval allows attentional priorities to be rapidly adjusted in a context-appropriate fashion. How the learning and reinstatement process works however, remains largely unclear.

The current study addresses two related issues: First, how is attentional control resolved when multiple contextual features could be used as cues? And second, what determines the expression and/or suppression of previously learned attentional associations?

Real-world environments are complex and contain many potential cues for directing attention. For instance, a coffee shop is a rich environment with numerous context cues (e.g., pictures on the wall, chairs, coffee, laptop), each of which someone could have different experiences with and afford different learned associations. A laptop for example, could cue attentional processing helpful for writing while other features of the coffee shop might cue attentional processing related to other experiences like socializing and people-watching. Given that all of these cues are available to set attentional priorities, what determines which associative relationships will be used to guide attention? On the one hand, context-based retrieval of attentional priorities might be completely obligatory: the mere presence of learned cues might trigger attention in a ballistic manner (e.g., Schneider \& Shiffrin, 1977; Shiffrin \& Schneider, 1977). Attention in this case would be at the mercy of the environment and attention is guided by the strongest association. On the other hand, our current task-goals might constrain the expression of previously learned associations. If someone sits down to write in the coffee shop, their current goals may allow the laptop to take priority over other environmental cues that are irrelevant to the task at hand.

Historically, perspectives of automaticity and attentional control have emphasized the role of practice and likened the learning of attentional associations to the development of automaticity in skill-learning (Logan, 1988, 1992; Schneider \& Shiffrin, 1977; Shiffrin \& Schneider, 1977). Under this view, forming an association requires many repeated experiences pairing a cue with a set of attentional priorities. Once formed however, it becomes habitual or hard-wired and the mere presence of the cue is sufficient to automatically adjust attention. Learned associations are therefore ballistic, difficult to suppress, and difficult to unlearn (e.g., Schneider \& Shiffrin, 1977; Posner \& Snyder, 1975a, 1975b; Shiffrin \& Schneider, 1977; Treisman, 1960, 1964, 1969). This view is often referenced in modern work on contextual cuing where it is assumed that learned associations require practice (e.g., Lehle \& Hübner, 2008; Crump, 2016), form without awareness or intention, and are triggered automatically by the mere presence of the cue (Crump et al., 2006; Crump \& Logan, 2010; Crump, Vaquero, \& Milliken, 2008; King et al., 2012; Reuss, Desender, Kiesel, \& Kunde, 2014; though, see Brosowsky \& Crump, 2016).

An alternative view---the memory-based perspective---takes a more nuanced view (Abrahamse et al., 2016; Brosowsky \& Crump, 2018; Crump, 2016; Egner, 2014; Norman, 1969, 1976). Here, practice is only important to the extent that it improves memory encoding and retrieval. Learning an association can be thought of as the process of encoding experiences where attentional priorities and cues co-occur. The expression of previously learned associations however, is dictated by memory retrieval---a similarity-based, cue-driven process (e.g., Jacoby, 1978; Hintzman, 1984, 1986, 1988). Presumably, the features of our immediate environment automatically cue memory and reinstate the retrieved attentional priorities (Logan, 1988, 1992). If a few prior experiences are easily retrievable (i.e., memories are distinct or very similar to the retrieval cue), they could support the reinstatement of a learned attentional association despite very little practice (e.g., Brosowsky \& Crump, 2018). Conversely, if prior experiences are difficult to retrieve (i.e., memories are indistinct or low in similarity to the retrieval cue), then learned associations may not be expressed, even after much practice. To the extent that you could manipulate memory retrieval, you should also be able manipulate the expression of learned associations.

Here, we examine whether task-relevance influences the retrieval of previously used attentional priorities. We hypothesized that the cue-driven, memory-retrieval process might be constrained to the features of our environment that are relevant to our on-going task and goals. If this is the case, then a previously learned association should be easily suppressed if that feature was made irrelevant to the task because that feature would no longer contribute to memory-retrieval. Furthermore, it would suggest that the mere presence of a cue is not sufficient for automatically directing attention, but instead requires task-integration (for a similar perspective, see Hommel (2004); Hommel (2019)).

Recent work on contextual cueing suggests that task-relevance is important for establishing new context-attention associations. Crump, Vaquero, \& Milliken (2008) used a Stroop prime-probe task where the distractor word is presented first followed by a color patch probe. Critically, they used the shape of the color patch probe, either a square or a circle, as the contextual dimension; one associated with a high frequency of conflict, the other with a low frequency. In one experiment participants were made aware of the shape-proportion contingencies but given no additional instructions. In the second, participants were given a secondary task to count the number of trials that contained a square. The first experiment failed to find evidence for context-specificity, whereas the second did. Although not testing task-relevance directly, Cañadas, Rodríguez-Bailón, Milliken, and Lupiáñez (2013) found similar effects. In this study they used images of male and female faces as context cues in a flanker task. When given instructions to think of the faces as members of the gender categories they found CSPC effects, but when given instructions to think of the faces as individuals they did not. These results suggest that making a context dimension relevant to the on-going task is important for establishing new associations and producing CSPC effects. However, the extent to which task-relevance influences the expression of previously learned associations is still unknown.

\hypertarget{experiment-1}{%
\section{Experiment 1}\label{experiment-1}}

Prior work suggests that the task-relevance of contextual information could be important for establishing new associations between context cues and attentional priorities. Simply put, CSPC effects tend to emerge when context cues are somehow made relevant to the task, but do not when made irrelevant (Cañadas et al., 2013; Crump et al., 2006, 2008). However, what determines the expression of previously learned context-attention associations remains unclear. In the current study, we examined whether manipulating task-relevance would allow participants to suppress a previously learned association and adopt a new context-specific control strategy.

@@

The critical measure of interest is the congruency effect produced in the frequency unbiased context. If, on the one hand, task-relevance has no impact on the use of context-attention associations we would expect no differences between the congruency effects when the high conflict context dimension is made task-relevant or the low conflict context dimension is made task-relevant. This could occur because all three contexts are treated as individual contexts throughout the whole experiment (e.g., Cañadas et al., 2013). Or it could occur because associations formed in the first block interfere with learning new associations in the second (e.g., Brosowsky \& Crump, 2016). On other hand, task-relevance may dictate the expression of previously learned associations. In this case, we would expect smaller congruency effects when the high conflict context dimension is made task-relevant as compared to when the low conflict context dimension is made task-relevant. This would demonstrate that participants were able to suppress a previously learned association and adopt a new context-specific control strategy.

@@

Experiment 1 used a 2x2x3 mixed design with task-relevant context (high conflict and low conflict) and unbiased-item congruency (congruent and incongruent) as within-subject factors, and context-type (object, social, and social/non-repeating) as the between-subjects factor.

All three conditions were constructed using the same general method. The experiment was divided into two phases. Each phase consisted of 144 flanker trials (48 trials per context), and 13 count response trials for a total of 314 trials. On the count response trials, participants indicated how many trials they had counted until that point. The count response trials occurred once for every 12 flanker trials and was randomly inserted between trial 6 and 12 of each 12-trial block. Each phase ended with one additional count response trial.

@@

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/figure2-1.pdf}
\caption{\label{fig:figure2}Results from Experiment 1 showing congruency effects (incongruent - congruent) for frequency unbiased items in reaction times (left) and error rates (right) as a function of the task-relevant context (100\% proportion congruent versus 0\% proportion congruent). Task-relevance was manipulated by the secondary feature-counting task.}
\end{figure}

\hypertarget{data-analysis}{%
\subsection{Data Analysis}\label{data-analysis}}

All analyses in this and the following experiments were performed in R (R Core Team, 2019) using a variety of R packages and resources (Aust \& Barth, 2018; Wickham, 2016; Wickham, François, Henry, \& Müller, 2019; Wilke, 2019). Of particular note: we used the afex and car packages to perform all the Null Hypothesis Significance Tests (Fox \& Weisberg, 2019; Singmann, Bolker, Westfall, \& Aust, 2019), the BayesFactor package to perform Bayesian analyses (Morey \& Rouder, 2018), and the conflictPower package to perform the simulation-based sensitivity and power analyses (Crump \& Brosowsky, 2019). All data and scripts can be found on the Open Science Framework (\url{https://osf.io/ztcyb/}).

We supplemented each Null Hypothesis Significance Test with a Bayes Factor analysis (Rouder, Speckman, Sun, Morey, \& Iverson, 2009). Using conventional frequentist testing it is not possible to quantify the evidence for a null effect. A Bayes Factor, however, is a continuous measure of the relative strength of evidence and can quantify the degree to which the data are compatible with the null over the alternative hypothesis (Dienes, 2014; Rouder et al., 2009). All Bayesian analyses were performed using the R package BayesFactor and Bayes Factors (\(\mathrm{BFs}\)) were calculated using its default settings (Morey \& Rouder, 2018). \(\mathrm{BF}_{\textrm{10}}\) indicates evidence in favor of the alternative hypothesis whereas \(\mathrm{BF}_{\textrm{01}}\) indicates evidence in favor of the null hypothesis. To simplify the interpretation, we report the Bayes Factor in the direction the data supports (e.g., \(\mathrm{BF}_{\textrm{01}}\) when there is more evidence in favor of the null over alternative hypothesis). As per previous recommendations, we refer to a \(\mathrm{BF > 3}\) as \enquote{moderate} and \(\mathrm{BF > 10}\) as \enquote{strong} evidence (Jeffreys, 1961; Rouder et al., 2009).

@@

The primary question of interest was whether the task-relevance of context features associated with different levels of conflict would influence the size of the congruency effect for frequency unbiased items (see Table 1). To that end, congruency effects (incongruent minus congruent performance) from frequency unbiased trials were submitted to a mixed analysis of variance (ANOVA) with task-relevant context (0\% and 100\% PC) as the within-subjects factor and context-type (object, social, and social/non-repeating) as the between-subjects factor. We supplement each Null Hypothesis Significance Test with a corresponding Bayesian analysis.

The results of the reaction time analysis revealed a significant effect of the task-relevant context, \(F(1, 128) = 8.30\), \(\mathit{MSE} = 3,217.08\), \(p = .005\), \(\hat{\eta}^2_p = .061\), 90\% CI \([0.01\), \(0.14]\), with moderate evidence in favor of the alternative, \(\mathrm{BF}_{\textrm{10}} = 6.87\) \([\pm 1.41\%]\). Specifically, we found smaller congruency effects when the context dimension associated with high conflict was made task-relevant as compared to when the low conflict dimension was made task-relevant. The main effect of context-type, however, was non-significant, \(F(2, 128) = 1.51\), \(\mathit{MSE} = 6,368.18\), \(p = .225\), \(\hat{\eta}^2_p = .023\), 90\% CI \([0\), \(0.07]\), with moderate evidence in favor of the null hypothesis, \(\mathrm{BF}_{\textrm{01}} = 3.96\) \([\pm 1.52\%]\). The two-way interaction between the task-relevant context and context-type was also non-significant, \(F(2, 128) < 0.01\), \(\mathit{MSE} = 3,217.08\), \(p = .996\), \(\hat{\eta}^2_p < .001\), 90\% CI \([0\), \(1]\), with moderate evidence in favor of the null hypothesis, \(\mathrm{BF}_{\textrm{01}} = 7.9\) \([\pm 1.49\%]\). Finally, the preferred model containing only the task-relevant context was preferred over the model containing the both factors and the interaction by a factor of \(54.25\) \([\pm 2.05\%]\). Therefore, congruency effects did not vary across context-types and, more importantly, the effect of task-relevance did not vary across context-types.

Similarly, the corresponding error analysis also resulted a significant effect of the task-relevant context, \(F(1, 128) = 6.72\), \(\mathit{MSE} = 27.87\), \(p = .011\), \(\hat{\eta}^2_p = .050\), 90\% CI \([0.01\), \(0.12]\), with moderate evidence in favor of the alternative evidence \(\mathrm{BF}_{\textrm{10}} = 3.13\) \([\pm 5.2\%]\); A non-significant effect of context-type, \(F(2, 128) = 0.81\), \(\mathit{MSE} = 51.40\), \(p = .449\), \(\hat{\eta}^2_p = .012\), 90\% CI \([0\), \(0.05]\), with moderate evidence for the null hypothesis, \(\mathrm{BF}_{\textrm{01}} = 7.54\) \([\pm 1.22\%]\); And a non-significant interaction between task-relevant context and context-type,\(F(2, 128) = 0.92\), \(\mathit{MSE} = 27.87\), \(p = .400\), \(\hat{\eta}^2_p = .014\), 90\% CI \([0\), \(0.05]\), with strong evidence in favor of the null hypothesis,\(\mathrm{BF}_{\textrm{01}} = 15.04\) \([\pm 2.57\%]\). Therefore, the congruency effects in error rates were also smaller when the context dimension associated with high conflict was made task-relevant and effects did not vary across context-type, corroborating the results of the reaction time analysis.

\begin{table}[htbp]
\caption{Reaction times and error rates from Experiment 1.}
\label{TR_table}
\centering
\begin{tabular}{lllcccc}
\toprule
 & & & \multicolumn{2}{c}{Congruent} & \multicolumn{2}{c}{Incongruent} \\
\cmidrule(rl){4-5}
\cmidrule(rl){6-7}
\multicolumn{1}{c}{Task-Relevant Context} & \multicolumn{1}{c}{Condition} & \multicolumn{1}{c}{PC} & \multicolumn{1}{c}{RT} & \multicolumn{1}{c}{ER} & \multicolumn{1}{c}{RT} & \multicolumn{1}{c}{ER}  \\
\midrule
\multirow{9}{*}{Low Conflict} & \multirow{3}{*}{Object} & \multicolumn{1}{l}{0\%} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{693 (26)} & \multicolumn{1}{l}{4.17 (0.73)} \\
& & \multicolumn{1}{l}{50\%} & \multicolumn{1}{l}{585 (23)} & \multicolumn{1}{l}{0.74 (0.27)} & \multicolumn{1}{l}{675 (28)} & \multicolumn{1}{l}{4.81 (0.82)} \\
& & \multicolumn{1}{l}{100\%} & \multicolumn{1}{l}{587 (25)} & \multicolumn{1}{l}{0.69 (0.23)} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{-} \\
\cmidrule(rl){2-7}
& \multirow{3}{*}{Social} & \multicolumn{1}{l}{0\%} & \multicolumn{1}{c}{-} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{745 (25)} & \multicolumn{1}{l}{5.86 (0.77)} \\
& & \multicolumn{1}{l}{50\%} & \multicolumn{1}{l}{612 (21)} & \multicolumn{1}{l}{0.39 (0.19)} & \multicolumn{1}{l}{722 (26)} & \multicolumn{1}{l}{6.78 (1.04)} \\
& & \multicolumn{1}{l}{100\%} & \multicolumn{1}{l}{616 (20)} & \multicolumn{1}{l}{0.63 (0.2)} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{-} \\
\cmidrule(rl){2-7}
& \multirow{3}{*}{Social (NR)} & \multicolumn{1}{l}{0\%} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{813 (25)} & \multicolumn{1}{l}{5.04 (0.74)} \\
& & \multicolumn{1}{l}{50\%} & \multicolumn{1}{l}{689 (23)} & \multicolumn{1}{l}{0.19 (0.14)} & \multicolumn{1}{l}{793 (27)} & \multicolumn{1}{l}{5.72 (1.06)} \\
& & \multicolumn{1}{l}{100\%} & \multicolumn{1}{l}{689 (25)} & \multicolumn{1}{l}{0.48 (0.17)} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{-} \\
\midrule
\multirow{9}{*}{High Conflict} & \multirow{3}{*}{Object} & \multicolumn{1}{l}{0\%} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{649 (16)} & \multicolumn{1}{l}{3.84 (0.6)} \\
& & \multicolumn{1}{l}{50\%} & \multicolumn{1}{l}{587 (17)} & \multicolumn{1}{l}{0.56 (0.21)} & \multicolumn{1}{l}{657 (17)} & \multicolumn{1}{l}{4.17 (1.02)} \\
& & \multicolumn{1}{l}{100\%} & \multicolumn{1}{l}{634 (18)} & \multicolumn{1}{l}{2.69 (0.58)} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{-} \\
\cmidrule(rl){2-7}
& \multirow{3}{*}{Social} & \multicolumn{1}{l}{0\%} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{722 (23)} & \multicolumn{1}{l}{5.77 (0.92)} \\
& & \multicolumn{1}{l}{50\%} & \multicolumn{1}{l}{633 (20)} & \multicolumn{1}{l}{0.78 (0.5)} & \multicolumn{1}{l}{722 (25)} & \multicolumn{1}{l}{4.75 (0.9)} \\
& & \multicolumn{1}{l}{100\%} & \multicolumn{1}{l}{673 (21)} & \multicolumn{1}{l}{1.84 (0.47)} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{-} \\
\cmidrule(rl){2-7}
& \multirow{3}{*}{Social (NR)} & \multicolumn{1}{l}{0\%} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{807 (26)} & \multicolumn{1}{l}{4.55 (0.8)} \\
& & \multicolumn{1}{l}{50\%} & \multicolumn{1}{l}{735 (28)} & \multicolumn{1}{l}{0.87 (0.36)} & \multicolumn{1}{l}{818 (30)} & \multicolumn{1}{l}{4.26 (0.8)} \\
& & \multicolumn{1}{l}{100\%} & \multicolumn{1}{l}{779 (34)} & \multicolumn{1}{l}{2.03 (0.4)} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{-} \\
 & & & & & & \\
\bottomrule
\multicolumn{7}{l}{\textit{Note}: RT = Reaction Time (ms);  ER = Error Rates (\%); PC = Proportion Congruent;} \\
\multicolumn{7}{l}{NR = Non-Repeating; Standard Errors are presented in parantheses.} \\
\end{tabular}%
\end{table}

\hypertarget{exploratory-analyses-order-effects}{%
\subsection{Exploratory analyses: Order effects}\label{exploratory-analyses-order-effects}}

Previous work has shown that shifting between list-wide proportion congruent phases produced asymmetrical effects. Abrahamse, Duthoo, Notebaert, and Risko (2013) found that shifting from a mostly-congruent to mostly-incongruent list produced a large change in the size of the congruency effect, whereas shifting from a mostly-incongruent to mostly-congruent list produced a relatively small change in congruency effects. The authors posited a learned attentional modulation in the mostly-incongruent block had carried over into the mostly-congruent block reducing what, typically, would be a large congruency effect. In our study, we were primarily interested in whether (a) participants could use context cues despite the presence of competing, contradictory cues, and (b) whether participants could switch from the use of one context cue to another. We therefore randomized the order across participants to control for the effects of order. However, given that we did find systematic shifts in the CSPC effect for unbiased items, we can also ask whether the shift between phases was asymmetrical.

To address this question, we collapsed over context-type and categorized participants according to the order of the counting task: 100\% congruent to 0\% congruent (100\%-0\%) or 0\% congruent to 100\% congruent (0\%-100\%). We again removed participants with error rates greater than 25\% and adopted the same outlier removal procedure as described above. We submitted congruency effects for frequency unbiased items to a 2x2 mixed ANOVA with task-relevant context (0\% and 100\% PC) as within-subject factors and phase order (100\%-0\% and 0\%-100\%) as the between-subjects factor, and supplemented each test with a corresponding Bayesian analysis.
